Debunking Neural Essay Scoring

### 2018-07-29 paper/acl17-latex/paper.tex
Modified Related Works section

Essay writing is usually a part of the student assessment  process. Several  organizations, such as Educational Testing Service (ETS)\cite{ets}, evaluate the writing skills of students in their examinations. Because of the large number of students participating in these exams, grading all essays is very time-consuming. 
Thus, automated essay scoring systems(AES) are used in evaluating and scoring student essays written based on a given prompt. The performance of these systems is assessed by comparing their scores assigned to a set of essays to human-assigned gold-standard scores.
Since the output of AES systems is usually a real valued number, the task is often addressed as a supervised machine learning task.
Recently, there is a lot of development of deep neural models for the task of automated essay scoring.
These models have shown to out perform the traditional models based on hand crafted features and currently have the state of art performance in this task. The current state of art model that we presently know of is from Taghipour et al. \cite{nea} and the state of art non-neural model is Enhanced AI Scoring Engine (EASE) \cite{ease}.
In this paper we explore the performance of deep-learning based systems on real life scenarios. We try to see if their performance is actually better than traditional non-neural based models in a qualitative sense and analyze the issues that these systems can cause.

@@

There exist many automated essay scoring systems \cite{shermis2013handbook} and some of them are being used in high-stakes assessments. E-rater \cite{attali2004automated} and Intelligent Essay Assessor \cite{foltz1999intelligent} are two notable examples of AES systems. In 2012, a competition on automated essay scoring called \textit{Automated Student Assessment Prize (ASAP)} was organized by Kaggle and sponsored by the Hewlett Foundation. A comprehensive comparison of AES systems was made in the ASAP competition. Many AES systems have been developed to date, and most of them have been built with hand-crafted features and supervised machine learning algorithms. Researchers have devoted a substantial amount of effort to design effective features for automated essay scoring. These features can be as simple as essay length \cite{chen2013automated} or more complicated such as lexical complexity, grammaticality of a text \cite{attali2004automated}, or syntactic features \cite{chen2013automated}. Readability features \cite{zesch2015task} have also been proposed in the literature as another source of information. Moreover, text coherence has also been exploited to assess the flow of information and argumentation of an essay \cite{chen2013automated}. A detailed overview of the features used in AES systems can be found in \cite{zesch2015task}.

Recently due to surge in deep-learning, a lot of neural models have been proposed which claim to beat the human level performance on evaluation metrics. Unlike traditional systems these systems, accepts an essay text as input directly and learns the features automatically from the data. A lot of models with varying architectures have been developed in the recent years. Some of them use a long short-term memory network (LSTM) \cite{taghipour2016neural}, while others use a convolutional neural networks (CNN) \cite{dong2016automatic} for the effect of automatically learning features. Some of them also use an augmented method to learn word embeddings \cite{alikaniotis2016automatic}.

@@

\subsection{Non-Neural Model}

The state of the art, open-source non neural model is Enhanced AI Scoring Engine (EASE) \cite{ease}. Unlike, the neural model, this model uses a lot of hand-crafted features to learn a vector representation of the input essay. These hand-crafted features encode the essay length, spelling mistakes, grammatical correctness and some of content-based features. Finally, a Gradient Boosting Regressor is used to predict the score of the essay.

@@

\section{Conclusions}
In this paper we have analyzed the behavior of neural model for automated essay scoring on various practical scenarios. We performed various experiments for effect of length, grammar, spelling, organization and found that neural models perform very poorly in these qualitative metrics even while giving state of art performance on the quantitative benchmarks. This shows that the problem of automated essay scoring is far from being solved and getting human level performance on benchmarks is not the solution. This also shows that the current benchmarks are not sufficient for evaluation and thus call for the need of better metrics. We have also suggested some solutions to problems like length-score correlation, misspelled words which seem to provide some robustness to above problems.

>>>>

Essay writing is one of the most popular ways to assess the skill of a student. Educational Testing Service (ETS)\cite{ets} uses essay scoring to assess the writing skill of students, but due to large number of students participating in these exams, it becomes very time consuming.
Automated essay scoring systems (AES) aims to automate the process of scoring essays. This is usually addressed as a supervised regression problem and the performance is judged by comparing the scores to a manually curated dataset of scored essays.
Recently, a lot of neural methods were proposed for this task and these have shown to give near human level performance in it. They seem to outperform the traditional feature based non-neural models like Enhanced AI Scoring Engine (EASE) \cite{ease}.
In this paper we explore the performance of deep-learning based systems on real life scenarios. We try to see compare their performance to traditional non-neural models in a qualitative sense and analyze the issues caused by them.

% Essay writing is usually a part of the student assessment  process. Several  organizations, such as Educational Testing Service (ETS)\cite{ets}, evaluate the writing skills of students in their examinations. Because of the large number of students participating in these exams, grading all essays is very time-consuming. 
% Thus, automated essay scoring systems(AES) are used in evaluating and scoring student essays written based on a given prompt. The performance of these systems is assessed by comparing their scores assigned to a set of essays to human-assigned gold-standard scores.
% Since the output of AES systems is usually a real valued number, the task is often addressed as a supervised machine learning task.
% Recently, there is a lot of development of deep neural models for the task of automated essay scoring.
% These models have shown to out perform the traditional models based on hand crafted features and currently have the state of art performance in this task. The current state of art model that we presently know of is from Taghipour et al. \cite{nea} and the state of art non-neural model is Enhanced AI Scoring Engine (EASE) \cite{ease}.
% In this paper we explore the performance of deep-learning based systems on real life scenarios. We try to see if their performance is actually better than traditional non-neural based models in a qualitative sense and analyze the issues that these systems can cause.

@@

Currently, there are a lot of essay scoring systems like E-rater \cite{attali2004automated}. The non-neural essay scorer that we use in our analysis is Enhanced AI Scoring Engine (EASE) \cite{ease}
by edx, which came 3rd in the kaggle competition named \textit{Automated Student Assessment Prize (ASAP)} sponsored by Hewlett Foundation.\\
Traditionally, the models used various kinds of hand engineered features like: grammaticality \cite{attali2004automated}, essay length \cite{chen2013automated}, text coherence and organization of ideas\cite{chen2013automated} etc.\\

Recently, due to surge in the area of deep-learning, a lot of neural models have been proposed which claim to beat the human level performance on evaluation metrics. These models accept an essay text as input and learn the features automatically. Various architectures using LSTMs \cite{taghipour2016neural}, CNNs\cite{dong2016automatic} etc. have been proposed for automatically learning the features.

% There exist many automated essay scoring systems \cite{shermis2013handbook} and some of them are being used in high-stakes assessments. E-rater \cite{attali2004automated} and E-rater \cite{attali2004automated} are two notable examples of AES systems. In 2012, a competition on automated essay scoring called \textit{Automated Student Assessment Prize (ASAP)} was organized by Kaggle and sponsored by the Hewlett Foundation. A comprehensive comparison of AES systems was made in the ASAP competition. Many AES systems have been developed to date, and most of them have been built with hand-crafted features and supervised machine learning algorithms. Researchers have devoted a substantial amount of effort to design effective features for automated essay scoring. These features can be as simple as essay length \cite{chen2013automated} or more complicated such as lexical complexity, grammaticality of a text \cite{attali2004automated}, or syntactic features \cite{chen2013automated}. Readability features \cite{zesch2015task} have also been proposed in the literature as another source of information. Moreover, text coherence has also been exploited to assess the flow of information and argumentation of an essay \cite{chen2013automated}. A detailed overview of the features used in AES systems can be found in \cite{zesch2015task}.

% Recently due to surge in deep-learning, a lot of neural models have been proposed which claim to beat the human level performance on evaluation metrics. Unlike traditional systems these systems, accepts an essay text as input directly and learns the features automatically from the data. A lot of models with varying architectures have been developed in the recent years. Some of them use a long short-term memory network (LSTM) \cite{taghipour2016neural}, while others use a convolutional neural networks (CNN) \cite{dong2016automatic} for the effect of automatically learning features. Some of them also use an augmented method to learn word embeddings \cite{alikaniotis2016automatic}.

@@

\subsection{Non-Neural Model}
This model numerous features to create a vector representation of the essay. The features are chosen to take essay length, spelling mistakes, grammatical correctness and some of content-based features in account thus giving potential to the model to perform well in scoring task.

@@

\section{Conclusions}
In this paper we have analyzed the behavior of neural model for automated essay scoring on various practical scenarios. We performed various experiments for effect of length, grammar, spelling, organization and found that neural models perform very poorly in these qualitative metrics even while giving state of art performance on the quantitative benchmarks. This shows that the problem of automated essay scoring is far from being solved and getting human level performance on benchmarks is not the solution. This also shows that the current benchmarks are not sufficient for evaluation and thus the need of better metrics. We have also suggested some solutions to problems like length-score correlation, misspelled words which seem to provide some robustness to above problems.
